---
title: "coding sample (Intermediate R)"
author: "Gloria Xiang"
output:
  pdf_document: default
  html_document: default
---

```{r load_packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(dslabs)
library(stats)
library(tibble)
library(purrr)
```

```{r}
ProjectTycho_Level1_v1_0_0 <- read_csv("ProjectTycho_Level1_v1.0.0.csv")
ProjectTycho_Level1<-unique(ProjectTycho_Level1_v1_0_0)
dim(ProjectTycho_Level1)
```

### For data frame `ProjectTycho_Level1`, drop its rows that have `disease = "DIPHTHERIA"`. After that, output the distinct values of `ProjectTycho_Level1$disease`.

```{r}
ProjectTycho_Level1<-ProjectTycho_Level1 %>% filter(disease!="DIPHTHERIA")
unique(ProjectTycho_Level1$disease)  ## only 7 distinct values
```

### Separate the column `epi_week` of `ProjectTycho_Level1` into two new columns named as `year` and `week` which are in the `integer` type. 
```{r Q3}
ProjectTycho_Level1<-ProjectTycho_Level1 %>% 
  separate(epi_week, into=c("year", "week"),sep=4, convert = TRUE)
head(ProjectTycho_Level1)
dim(ProjectTycho_Level1)
```

### From `ProjectTycho_Level1`, create a new data frame, named as `ProjectTycho_count`, that contains the count of cases of each disease for each state at each year, with column names `disease`, `state`, `year`, `weeks_reporting` and `count`. 
```{r Q4}
ProjectTycho_count<-ProjectTycho_Level1 %>% 
  drop_na(cases)  %>% 
  group_by(disease, loc, year)  %>% 
  summarize(week_reporting=n(),count=sum(cases))   %>% 
  rename(state=loc) %>% 
  ungroup()
head(ProjectTycho_count)
dim(ProjectTycho_count)
```



```{r}
region_state <- list(
  Northeast = state.name[state.region=="Northeast"], 
  South = state.name[state.region=="South"],
  North_Central = state.name[state.region=="North Central"], 
  West = state.name[state.region=="West"] 
  ) 
str(region_state)
```



### Add `"District Of Columbia"` into the element `South` of the list `region_state` . Output the structure of the new `region_state` list. 

```{r}
region_state$South <- c(region_state$South, "District Of Columbia")
str(region_state) # the element South must be shown with chr [1:17] not chr [1:16]
```



###  From the data `us_contagious_diseases` of package `dslabs`, ignoring the variable `weeks_reporting`, compute the yearly incidence rate of each disease for each of the 4 geographic regions (Northeast, South, North Central, West). Store the result into a new data frame, named as `region_incidence`, with columns `disease`, `region`, `year`, and `incidence_per_millon` (i.e., the yearly incidence rate times one million). 
```{r}
data(us_contagious_diseases)
region_incidence <- us_contagious_diseases %>%
  select(-weeks_reporting) %>% 
  drop_na() %>%
  mutate(region = fct_collapse(state,
    Northeast = region_state$Northeast,
    South = region_state$South,
    North_Central = region_state$North_Central,
    West = region_state$West),
    .after = disease) %>%
  group_by(disease, region, year) %>%
  summarize(incidence_per_millon = sum(count)/sum(population)*1e6) %>%
  ungroup()
head(region_incidence)
dim(region_incidence)
```



###  In the data frame `region_incidence`, choose an appropriate `str_` function to add the suffix `_US` to the values in the column `region`. Make sure that the resulting column `region` is a factor with levels in the order: `North_Central_US`, `Northeast_US`, `South_US`, `West_US`.  
```{r}
region_incidence <- region_incidence %>% 
  mutate(region = factor(str_c(region, "_US"))) 
head(region_incidence) # the column region must be shown with <fct> below the column name
levels(region_incidence$region) # must be in the requested order
```



###  From `region_incidence`, use `ggplot` to provide a line graph that shows the trend of the yearly incidence of disease `Hepatitis A` for all the 4 geographic regions. You need to appropriately reorder the factor levels of `region` to make the graph easier to read with the legend. 

```{r}
region_incidence.Hepatitis_A <- region_incidence %>% filter(disease == "Hepatitis A")
region_incidence.Hepatitis_A %>% 
  ggplot(aes(
    x = year, y = incidence_per_millon,
    colour = region
  )) +
  # geom_point() +
  geom_line() +
  labs(colour = "Region") +
  xlab("Year") +
  ylab("Incidence per millon")
```

It may be better to make the line colors at the year 1970 (around the peaks) line up with the legend:
```{r}
region_incidence.Hepatitis_A %>% 
  ggplot(aes(
    x = year, y = incidence_per_millon,
    colour = fct_relevel(region, "West_US","Northeast_US", "South_US", "North_Central_US")
  )) +
  # geom_point() +
  geom_line() +
  labs(colour = "Region") +
  xlab("Year") +
  ylab("Incidence per millon")
```


